---
title: "Homework 6"
author: "Juan Cambeiro"
date: "2022-12-03"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv) 
library(broom) 

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

To obtain a distribution for $\hat{r}^2$, we'll follow basically the same procedure we used for regression coefficients: draw bootstrap samples; the a model to each; extract the value I'm concerned with; and summarize. Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r weather_df, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

In this example, the $\hat{r}^2$ value is high, and the upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for $R^2$, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, because the shape isn't symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.

We can produce a distribution for $\log(\beta_0 * \beta1)$ using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

As with $r^2$, this distribution is somewhat skewed and has some outliers. 

The point of this is not to say you should always use the bootstrap -- it's possible to establish "large sample" distributions for strange parameters / values / summaries in a lot of cases, and those are great to have. But it is helpful to know that there's a way to do inference even in tough cases. 

### Problem 2

First, I will load and clean the homicide data that was gathered by The Washington Post and describe the raw data.
```{r}
homicide =
  read_csv("data/homicide-data.csv") %>%
  janitor::clean_names()
```

**Description of raw data:** There are now `r nrow(homicide)` observations and `r ncol(homicide)` variables in the dataset.


Next, I will tidy the data by using `mutate` to make a `city_state` variable and I will also make a binary `case_solved` variable to indicate whether a homicide is resolved (1 for solved, 0 for not solved). I also ensure appropriate data types are used. I use `filter` to omit the following cities: Dallas, TX; Phoenix, AZ; Kansas City, MO; Tulsa, AL. I also use `filter` to limit the analysis to those victim_race is White or Black.
```{r}
homicide_tidy = homicide %>%
  mutate(city_state = as.factor(str_c(city, state, sep = ", ")), 
         resolved = if_else(disposition == "Closed by arrest", 1, 0),
         reported_date = as.Date(as.character(reported_date),"%Y%m%d"), 
         victim_age = as.numeric(victim_age), 
         victim_sex = as.factor(victim_sex),
         victim_race = as.factor(victim_race), 
         city = as.factor(city),
         state = as.factor(state)) %>%
  filter(city_state != "Dallas, TX",
         city_state != "Phoenix, AZ", 
         city_state != "Kansas City,MO", 
         city_state != "Tulsa, AL") %>% 
  filter(victim_race == "White" | victim_race == "Black")
```

**Description of tidied data:** There are now `r nrow(homicide_tidy)` observations and `r ncol(homicide_tidy)` variables in the dataset.


Now, I use the `glm` function to fit a logistic regression for Baltimore, MD with the binary `resolved` variable as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. I save the output as an R object, apply `broom::tidy`, and obtain the estimate and confidence interval of the adjusted OR for homicides resolved comparing male victims to female victims.

```{r}
baltimore_glm =
  homicide_tidy %>%
  filter(city_state == "Baltimore, MD")
output = glm(resolved ~ victim_age + victim_sex + victim_race, data = baltimore_glm, family = "binomial") %>% 
  tidy(exponentiate = 1, conf.int = 1, conf.level = 0.95) %>%
  filter(term == "victim_sexMale") %>%
  select(term, estimate, conf.low, conf.high)
output
```

**Adjusted OR for solving homicides in Baltimore, comparing male victims to female victims:** 0.426 (95% CI: 0.324, 0.558)

Next, I run `glm` for each of the cities in my dataset to extract the adjusted ORs and corresponding CIs for solving homicides comparing male victims and female victims. I do this using  `purrr::map`, list columns, and `unnest` to create a dataframe with estimated ORs and CIs for each city. 

```{r}
all_cities_glm =
  homicide_tidy %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = map(data, ~glm(resolved ~ victim_age + victim_sex + victim_race, family = "binomial", data = .x)),
    results = map(models, tidy, exponentiate = 1, conf.int = 1, conf.level = 0.95)) %>%
  select(-data, -models) %>% 
  unnest(results) %>%
  filter(term == "victim_sexMale") %>%
  select(city_state, term, estimate, conf.low, conf.high)
all_cities_glm
```

Now, I create a plot to show the estimated ORs and corresponding CIs by city, organized by estimated OR. 

```{r}
all_cities_glm %>%
  ggplot(aes(x = estimate, y = reorder(city_state, +estimate))) +
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
  labs(title = "Odds Ratios for Homicide Resolutions By City, Comparing Male to Female Victims",
    x = "OR",
    y = "City")
```


